\documentclass[12pt, a4paper]{article}

\usepackage{setspace, graphicx, lineno, caption, color, float}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amstext} %to enter text in mathematical formulae
\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{multirow} %connecting columns in tables
\usepackage{multicol}
\usepackage{longtable}

\usepackage[left=2.5cm,top=2.5cm,right=2.5cm, bottom=2.5cm,nohead]{geometry}
%paragraph formatting
\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm}

\setcounter{table}{1}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}
\section*{Appendix 4: Action space}
The sub-actions link management decisions to the population of \textit{Alopecurus myosuroides}, and thus the yield of wheat and so the reward function. Each sub-action also has a cost which affects the reward function.   
\begin{longtable}[h]{c p{9cm} p{4cm}}
\caption{Management sub-actions and their effects on the population model\label{table:actions}}\\
	\hline
	\textbf{Sub-action} & \textbf{Effect on system model} & \textbf{Management parameters}\\
	\hline	
	\multicolumn{3}{l}{\textit{Crop choice}: $a_k$}\\
	wheat & higher survival rate for \textit{Alopecurus myosuroides} without selective herbicides. & Highest profit in the absence of \textit{Alopecurus myosuroides}. \\
	alt & \textit{Alopecurus myosuroides} survival reduced under alternative crop due to competition or broad spectrum herbicide use before sowing. We parameterize the alternative crop as spring barley, a common rotation used in the control of \textit{Alopecurus myosuroides}. & Lower income than winter wheat.\\
	fallow & no crop planted. All above ground plants killed as we assume used alongside plowing or broad spectrum herbicide. & No production, costs for killing above ground plants.\\
	\multicolumn{3}{l}{\textit{Herbicide}: $a_h$}\\
	no herb & No effect & No cost\\ 
	herb R & Individuals with resistance alleles $RR$ or $Rr$ are protected, but all other individuals have greatly reduced survival. & Application cost.\\
	herb Q & Individuals with resistance alleles $QQ$ or $Qq$ are protected, but all other individuals have greatly reduced survival. & Application cost.\\  
	both & Individuals with both resistance alleles ($RRQQ$, $RRQq$, $RrQQ$, $RrQq$) are protected, but all other individuals have greatly reduced survival. & Double application cost.\\  
	\multicolumn{3}{l}{\textit{Seed bank management}: $a_b$}\\
	1 & Moves seeds from one level of the seed bank to the other. & Fixed cost.\\
	0 & No effect & No cost.\\
	\multicolumn{3}{l}{\textit{spot control}: $a_s$}\\
	1 & No effect of above ground population. & No cost.\\
	0 & Kills all remaining above ground plants. & Cost scales with above ground post control populations. \\
	\hline
\end{longtable}

These sub-actions are combined to create a single action $a_j^t$ that could be taken in a time step (Table \ref{table:action_space}). However, some sub-action combinations do not make sense, for example applying herbicide to the population when $a_k =$ 'fallow', since we assume all above ground plants are destroyed under this crop choice. The list of all allowed sub-action combinations is the action space $\mathbf{A}$.   

\begin{longtable}[h]{c c c c c c}
\caption{Action space ($\mathbf{A}$) with all eligable combinations of sub actions\label{table:action_space}}\\
	\hline
	$\mathbf{a_j}$ & $\mathbf{a_k}$ & $\mathbf{a_h}$ & $\mathbf{a_b}$ & $\mathbf{a_s}$\\
	\hline
	$a_1$ & wheat & no herb & 0 & 1\\
	$a_2$ & wheat & no herb & 0 & 0\\
	$a_3$ & wheat & no herb & 1 & 1\\
	$a_4$ & wheat & no herb & 1 & 0\\
	$a_5$ & wheat & herb R & 0 & 1\\
	$a_6$ & wheat & herb R & 0 & 0\\
	$a_7$ & wheat & herb R & 1 & 1\\
	$a_8$ & wheat & herb R & 1 & 0\\
	$a_9$ & wheat & herb Q & 0 & 1\\
	$a_{10}$ & wheat & herb Q & 0 & 0\\
	$a_{11}$ & wheat & herb Q & 1 & 1\\
	$a_{12}$ & wheat & herb Q & 1 & 0\\
	$a_{13}$ & wheat & both & 0 & 1\\
	$a_{14}$ & wheat & both & 0 & 0\\
	$a_{15}$ & wheat & both & 1 & 1\\
	$a_{16}$ & wheat & both & 1 & 0\\
	$a_{17}$ & alt & no herb & 0 & 1\\
	$a_{18}$ & alt & no herb & 0 & 0\\
	$a_{19}$ & alt & no herb & 1 & 1\\
	$a_{20}$ & alt & no herb & 1 & 0\\
	$a_{21}$ & alt & herb R & 0 & 1\\
	$a_{22}$ & alt & herb R & 0 & 0\\
	$a_{23}$ & alt & herb R & 1 & 1\\
	$a_{24}$ & alt & herb R & 1 & 0\\
	$a_{25}$ & alt & herb Q & 0 & 1\\
	$a_{26}$ & alt & herb Q & 0 & 0\\
	$a_{27}$ & alt & herb Q & 1 & 1\\
	$a_{28}$ & alt & herb Q & 1 & 0\\
	$a_{29}$ & alt & both & 0 & 1\\
	$a_{30}$ & alt & both & 0 & 0\\
	$a_{31}$ & alt & both & 1 & 1\\
	$a_{32}$ & alt & both & 1 & 0\\
	$a_{33}$ & fallow & no herb & 0 & 1\\
	$a_{34}$ & fallow & no herb & 1 & 1\\
	\hline
\end{longtable}

\end{document}